---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
str(av)
```

## Death Mutate

```{r}
library(tidyverse)
deaths <- av %>%
  pivot_longer(
    cols = Death1:Death5,
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Death != "")
head(deaths, 10)
str(deaths)
```

## Returns Mutate

```{r}
library(tidyverse)

returns <- av %>%
  pivot_longer(
    cols = Return1:Return5,
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Return != "")

head(returns)
```

## Average Deaths

```{r}
total_avengers <- nrow(av)
total_deaths <- deaths %>% 
  filter(Death == "YES") %>% 
  nrow()

avg_deaths <- total_deaths / total_avengers
print(avg_deaths)
```

## Individually

### Harrison Voss

#### FiveThirtyEight statement

> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team.

#### Code

```{r}
total_avengers <- nrow(av)

avengers_who_died <- deaths |>
  filter(Death == "YES") |>
  distinct(Name.Alias) |>
  nrow()


print(avengers_who_died)

percent_died <- avengers_who_died / total_avengers

print(percent_died* 100)
```

#### Answer

It seems that the number of people who died was 64, not 69, and that the percentage is 37% not 40%.

This is not a very big discrepancy, but my findings do differ from the claim in the analysis.
