---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
str(av)
```

## Death Mutate

```{r}
library(tidyverse)
deaths <- av %>%
  pivot_longer(
    cols = Death1:Death5,
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Death != "")
head(deaths, 10)
str(deaths)
```

## Returns Mutate

```{r}
library(tidyverse)

returns <- av %>%
  pivot_longer(
    cols = Return1:Return5,
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Return != "")

head(returns)
```

## Average Deaths

```{r}
total_avengers <- nrow(av)
total_deaths <- deaths %>% 
  filter(Death == "YES") %>% 
  nrow()

avg_deaths <- total_deaths / total_avengers
print(avg_deaths)
```

## Individually

### Harrison Voss

#### FiveThirtyEight statement

> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team.

#### Code

```{r}
total_avengers <- nrow(av)

avengers_who_died <- deaths |>
  filter(Death == "YES") |>
  distinct(Name.Alias) |>
  nrow()


print(avengers_who_died)

percent_died <- avengers_who_died / total_avengers

print(percent_died* 100)
```

#### Answer

It seems that the number of people who died was 64, not 69, and that the percentage is 37% not 40%.

This is not a very big discrepancy, but my findings do differ from the claim in the analysis.

### Nhu Phan

#### Claim to check
> Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. In fact, Hawkeye died twice!

```{r}
library(tidyverse)

# Define the nine MCU Avengers mentioned in the article
mcu_avengers <- c(
  "Anthony Edward \"Tony\" Stark",   # Iron Man
  "Robert Bruce Banner",            # Hulk
  "Steven Rogers",                  # Captain America
  "Thor Odinson",                   # Thor
  "Clinton Francis Barton",                   # Hawkeye
  "Natalia Alianovna Romanova", # Black Widow
  "Wanda Maximoff",                 # Scarlet Witch
  "Pietro Maximoff",                # Quicksilver
  "Victor Shade (alias)"                          # Vision
)

mcu_deaths <- deaths %>%
  filter(Name.Alias %in% mcu_avengers & Death == "YES") %>%
  count(Name.Alias, name = "Times_Died") %>%
  arrange(desc(Times_Died))

mcu_deaths
```

#### Conclusion
Based on the data, every one of the nine MCU Avengers has died at least once in the comics, confirming the main claim from the article. However, the statement that only Hawkeye died twice is inaccurate. The dataset shows that several Avengers — including Iron Man, Black Widow, Hulk, Captain America, Vision, and others — have died more than once. Hawkeye does have the highest number of deaths (four), but he is not the only Avenger to experience multiple deaths. Therefore, while the overall idea that the Avengers frequently die and return is true, the specific detail about Hawkeye being unique in this regard is not supported by the data.

### Mahathi Reddy
> But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.
```{r}

library(dplyr)

# Load the data
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
```
```{r}
# Check the structure of death and return columns
av %>%
  select(URL, Death1, Return1, Death2, Return2, Death3, Return3, Death4, Return4, Death5, Return5) %>%
  head(10)
```

```{r}
# Fact-check: Return rate from FIRST death
first_death_return <- av %>%
  filter(Death1 == "YES") %>%  # Only those who died at least once
  summarise(
    total_first_deaths = n(),
    returned_from_first = sum(Return1 == "YES", na.rm = TRUE),
    return_rate_first = mean(Return1 == "YES", na.rm = TRUE),
    return_percentage_first = mean(Return1 == "YES", na.rm = TRUE) * 100
  )

cat("FIRST DEATH:\n")
print(first_death_return)
cat("\nReturn rate:", round(first_death_return$return_rate_first, 3), 
    "(or", round(first_death_return$return_percentage_first, 1), "%)\n")
cat("This is approximately", round(first_death_return$return_rate_first * 3, 1), 
    "in 3, or", round(first_death_return$return_rate_first, 2), "\n\n")
```

```{r}
# Fact-check: Return rate from SECOND death
second_death_return <- av %>%
  filter(Death2 == "YES") %>%  # Only those who died a second time
  summarise(
    total_second_deaths = n(),
    returned_from_second = sum(Return2 == "YES", na.rm = TRUE),
    return_rate_second = mean(Return2 == "YES", na.rm = TRUE),
    return_percentage_second = mean(Return2 == "YES", na.rm = TRUE) * 100
  )

cat("SECOND DEATH:\n")
print(second_death_return)
cat("\nReturn rate:", round(second_death_return$return_rate_second, 3), 
    "(or", round(second_death_return$return_percentage_second, 1), "%)\n\n")
```

```{r}
# Fact-check: Return rate from THIRD death
third_death_return <- av %>%
  filter(Death3 == "YES") %>%  # Only those who died a third time
  summarise(
    total_third_deaths = n(),
    returned_from_third = sum(Return3 == "YES", na.rm = TRUE),
    return_rate_third = mean(Return3 == "YES", na.rm = TRUE),
    return_percentage_third = mean(Return3 == "YES", na.rm = TRUE) * 100
  )

cat("THIRD DEATH:\n")
print(third_death_return)
cat("\nReturn rate:", round(third_death_return$return_rate_third, 3), 
    "(or", round(third_death_return$return_percentage_third, 1), "%)\n\n")
```

```{r}
# Combined second and third death return rate
second_and_third_combined <- av %>%
  filter(Death2 == "YES" | Death3 == "YES") %>%
  mutate(
    died_second_or_third = (Death2 == "YES") + (Death3 == "YES"),
    returned_second_or_third = (Return2 == "YES") + (Return3 == "YES")
  ) %>%
  summarise(
    total_deaths_2_or_3 = sum(died_second_or_third, na.rm = TRUE),
    total_returns_2_or_3 = sum(returned_second_or_third, na.rm = TRUE),
    combined_return_rate = total_returns_2_or_3 / total_deaths_2_or_3,
    combined_return_percentage = (total_returns_2_or_3 / total_deaths_2_or_3) * 100
  )

cat("SECOND OR THIRD DEATH COMBINED:\n")
print(second_and_third_combined)
cat("\nCombined return rate:", round(second_and_third_combined$combined_return_rate, 3), 
    "(or", round(second_and_third_combined$combined_return_percentage, 1), "%)\n\n")
```

```{r}
# Summary comparison table
summary_table <- tibble(
  Death_Number = c("First Death", "Second Death", "Third Death", "Second or Third Combined"),
  Total_Deaths = c(first_death_return$total_first_deaths,
                   second_death_return$total_second_deaths,
                   third_death_return$total_third_deaths,
                   second_and_third_combined$total_deaths_2_or_3),
  Returns = c(first_death_return$returned_from_first,
              second_death_return$returned_from_second,
              third_death_return$returned_from_third,
              second_and_third_combined$total_returns_2_or_3),
  Return_Rate = c(first_death_return$return_rate_first,
                  second_death_return$return_rate_second,
                  third_death_return$return_rate_third,
                  second_and_third_combined$combined_return_rate),
  Return_Percentage = c(first_death_return$return_percentage_first,
                        second_death_return$return_percentage_second,
                        third_death_return$return_percentage_third,
                        second_and_third_combined$combined_return_percentage)
)

cat("SUMMARY TABLE:\n")
print(summary_table)
```

```{r}
# Verify the "2-in-3" claim (which is approximately 0.667 or 66.7%)
cat("\n=== VERIFICATION ===\n")
cat("Claim: 2-in-3 chance (66.7%) of return from first death\n")
cat("Actual:", round(first_death_return$return_percentage_first, 1), "%\n")
cat("Difference:", round(first_death_return$return_percentage_first - 66.7, 1), "percentage points\n\n")

cat("Claim: 50% chance of return from second or third death\n")
cat("Actual:", round(second_and_third_combined$combined_return_percentage, 1), "%\n")
```

There is no difference between the claim and the returned value. This means that there is a 2 in 3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death. 
